---
title: "Greater North Sea mixoplankton"
subtitle: "A first analysis"
author: "Willem Stolte"
date: "17/06/2020"
output: html_notebook
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align="center")
require(sf)
require(tidyverse)
require(lubridate)
dataDir <- "../data/derived_data"
```

```{r styleDef, echo = T}
EMODnetMapCleanStyle <- theme_minimal() +
  theme(axis.text = element_blank(),
        axis.title = element_blank())

EMODnetMapNormalStyle <- 
  theme(axis.text = element_text(size = 12, color = "blue"),
        axis.title = element_text(size = 14, color = "blue"),
        legend.text = element_text(size = 10),
        legend.position = "right",
        legend.key.width = unit(10,"mm"),
        legend.key.height = unit(10, "mm"),
        plot.background = element_rect(fill = "grey90", colour = "white", size = 0),
        panel.border = element_blank()
)


```



# Introduction

## Documentation

Github project: https://github.com/wstolte/EMODnet-Biology-Phytoplankton-NorthSea

# Data extraction

## Temporal scale

Data from 1995 until now are considered for the current product.

## Geographical scale

The regions that were selected were assembled from the intersection of the IHO regions and the EEZ from the different countries. These subregions have ID's that can be used in the WFS query to the EMODnet Biology database.

```{r mapGreaterNorthSea, message=F, warning=F, comment=F, out.width="50%"}
regions <- sf::st_read(quiet = T, dsn = "../data/derived_data/simplified_greater_north_sea-selection_from_eez-iho_v4.geojson")
regions %>% ggplot() +
  geom_sf(fill = "blue", color = "white") +
  geom_sf_text(aes(label = mrgid), size = 2, color = "white") +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
```

## Strategy

For each of these subregions, data extraction was done in a two-step approach. 

1. Extract phytoplankton data by selecting only observations linkted to the trait "phytoplankton".
1. The unique datasets obtained in the first steps were inspected, and suitable datasets, which were expected to contain phytoplankton, were selected. The suitable datasets were then downloaded completely (without selection for traits)


## Extraction of complete relevant datasets

Data were extracted from EMODnet Biology via WFS, using the requestData.R script. These data are stored locally and taken into this script here. 


```{r, message=F, warning=F}
# all2Data <- read_delim(file.path(dataDir, "all2Data.csv"), delim = ";")
load(file.path(dataDir, "all2Data.Rdata"))
```

In total, this resulted in `r length(all2Data$id)` observations

Per subregion, the number of observations/km^2^ looks like this:

```{r, message=F, warning=F}
regionN <- all2Data %>%
  group_by(mrgid) %>% summarize(n = n()) %>% ungroup() %>%
  mutate(mrgid = as.numeric(mrgid))

regions %>% right_join(regionN, by = c(mrgid = "mrgid")) %>%
  mutate(n_km2 = n/area_km2) %>%
  ggplot() +
  geom_sf(aes(fill = n_km2)) +
  scale_fill_viridis_c()
```









 










